---
title: "Main Functions"
vignette: >
  %\VignetteIndexEntry{Main Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
format:
  html:
    toc: true
    embed-resources: true
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This vignette lays out the functions that `demogendr`
will eventually export. Each chunk below has a call and a short description of what it should do. As the
package grows, these calls should become runnable.

```{r}
library(demogendr)

democracy_data <- load_data()
```

## 1. `plot_women_in_parliament()` — line plot

```{r}
plot_women_in_parliament(
  democracy_data,
  countries = c("United States", "Sweden", "Honduras"),
  min_year = 1990,
  color_by = "is_democracy"
)
```

Should draw a `ggplot2` line chart of the proportion of parliamentary
seats held by women (`prop_in_par`) over time, with one line per
country (`United States, Sweeden, and Honduras`). The `color_by` argument lets the user shade each country's
line based on democratic status in that year, this should make it easy to see transitions in leadership.

## 2. `summarise_leader_gender_gap()` — summary table

```{r}
summarise_leader_gender_gap(
  democracy_data,
  group_var = "decade",
  leader = "president"
)
```

Should return a tidy summary tibble grouped by the requested variable
(`decade`, `year`, `is_democracy`, `region`, ...) with columns for the
proportion of female vs. male leaders and the mean accession age for each gender. Passing `leader = "monarch"`
should use the same calculations except for it should use monarchs instead of presidents.
This is similar to `summarise_leadership_gender_gap()` which we
wrote in Project 2 this would allow us to use it outside of a dashboard.

## 3. `top_transition_countries()` — ranking calculation

```{r}
top_transition_countries(
  democracy_data,
  start_year = 1950,
  end_year = 2020,
  n = 10
)
```

Should return the number of countries with the most regime transitions
(Both into and out of democracy) over an inputed number of years. The
function should compute `to_democracy` / `from_democracy` using lag and `is_democracy` within each country, then sum and sort. 
This is a similar to `topn_countries()` from the Project 2 dashboard, but we can use it on a subset of the dataset.
