---
title: "Main Functions"
vignette: >
  %\VignetteIndexEntry{Main Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
format:
  html:
    toc: true
    embed-resources: true
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.height = 4.5
)
```

`demogendr` bundles a country-year panel of democratic status, presidential
and monarchic leadership, and women's representation in parliament. This
vignette walks through one example use of each of the package's seven core
functions, starting from loading the data.

```{r}
#| label: load
library(demogendr)

democracy_data <- load_data()
dim(democracy_data)
```

## 1. `load_data()` — load the bundled dataset

`load_data()` reads the merged `Combined_Democracy_Data.csv` that ships in the
package's `inst/extdata/` folder and returns it as a tibble, one row per
country-year.

```{r}
#| label: load-preview
head(democracy_data[, c("country_name", "year", "is_democracy",
                        "is_presidential", "prop_in_par")])
```

## 2. `clean_data()` — cleaning / wrangling

`clean_data()` enriches the raw panel with derived columns: age-at-accession
for presidents and monarchs, a `has_female_leader` flag, a `decade` label, and
year-over-year democratic-transition indicators.

```{r}
#| label: clean
clean <- clean_data(democracy_data)

head(clean[, c("country_name", "year", "decade",
               "president_age_at_accession",
               "has_female_leader",
               "to_democracy", "from_democracy")])
```

## 3. `summarise_leader_gender_gap()` — summary table

Returns a tidy summary grouped by the requested variable (here `decade`) with
the proportion of female vs. male leaders and the mean accession age for each
gender. Passing `leader = "monarch"` runs the same calculation for monarchs.

```{r}
#| label: summarise-gap
summarise_leader_gender_gap(
  democracy_data,
  group_var = "decade",
  leader = "president"
)
```

## 4. `top_transition_countries()` — ranking calculation

Computes `to_democracy` / `from_democracy` events from `is_democracy` within
each country (using a `data.table` within-country lag), then sums and sorts to
return the countries with the most regime transitions over a chosen window.

```{r}
#| label: top-transitions
top_transition_countries(
  democracy_data,
  start_year = 1950,
  end_year = 2020,
  n = 10
)
```

## 5. `plot_women_in_parliament()` — line plot

Draws a `ggplot2` line chart of the proportion of parliamentary seats held by
women (`prop_in_par`) over time, one line per country, colored by democratic
status so leadership transitions are easy to see.

```{r}
#| label: plot-parliament
plot_women_in_parliament(
  democracy_data,
  countries = c("United States", "Sweden", "Honduras"),
  min_year = 1990,
  color_by = "is_democracy"
)
```

## 6. `plot_accession_ages()` — ribbon-and-line plot

Shows the average age at which presidents and monarchs took office, split by
gender, with a shaded ribbon highlighting the gender gap for each leader type.

```{r}
#| label: plot-ages
plot_accession_ages(
  democracy_data,
  min_year = 1980,
  max_year = 2020
)
```

## 7. `analyse_parliament_gap()` — statistical analysis

Runs a Welch two-sample t-test comparing women's parliamentary representation
across two groups. The grouping column is supplied unquoted via tidy
evaluation, here comparing democracies to non-democracies.

```{r}
#| label: analyse-gap
analyse_parliament_gap(democracy_data, group_col = is_democracy)
```

### `plot_regime()` — regime-share area chart

Shows the share of countries classified under a given regime category over
time, as a filled area-and-line chart.

```{r}
#| label: plot-regime
plot_regime(
  democracy_data,
  regime_type = "Parliamentary democracy",
  min_year = 1960,
  max_year = 2020
)
```

### `compare_regime_types2()` — t-test with string arguments

A string argument alternative to `analyse_parliament_gap()` that compares a
numeric outcome between two named regime values.

```{r}
#| label: compare-regime
compare_regime_types2(
  democracy_data,
  outcome_var = "prop_in_par",
  regime_var  = "is_democracy",
  regime1     = TRUE,
  regime2     = FALSE
)
```
